### 2022-09-19 AISTATS2023/paper.tex
added examples of 4.1.1

% If you use natbib package, activate the following three lines:
\usepackage[round]{natbib}
\renewcommand{\bibname}{References}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}
\bibliographystyle{plainnat}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}

% If you use BibTeX in apalike style, activate the following line:
%\bibliographystyle{apalike}

@@

\section{SYNTHETIC EXAMPLES}

>>>>

% If you use natbib package, activate the following three lines:
\usepackage[round]{natbib}
\renewcommand{\bibname}{References}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}
\bibliographystyle{plainnat}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{graphicx}
\graphicspath{ {./../examples/} }
% If you use BibTeX in apalike style, activate the following line:
%\bibliographystyle{apalike}

@@

\clearpage
\section{EXPERIMENTS}

The experimental section is divided in simulation and real-world
examples. The simulation examples, where the data-generating
distribution \(p(\mathbf{X})\) and the predictive function
\(f(\cdot)\) are defined by us, are used for comparing all methods
against a solid ground-truth. Having the freedom to experiment with
\(p(\mathbf{X}), f(\cdot)\), we explore the behavior of each method in
different settings for highlighting important aspects
(advantages/disadvantages) of each one. In particular, we design two
end-to-end simulation examples; At the first one, we focus on
comparing NAME method versus PDP-ICE in terms of quantifying the
heterogeneous effects for a range of different black-box functions. At
the second one, we provide empirical clues that NAME method achieves a
better approximation compared to traditional ALE, due to appropriate
bin-splitting. Afterwards, we also apply the methods on some
real-world datasets. In these cases, it is infeasible to obtain ground
truth information, therefore, we perform evaluation at two
levels. First, we compare visually NAME versus PDP with ICE
plots. Second, we hold as ground-truth the effects computed on the
full training-set and we, then, perform subsampling to evaluate the
robustness of the bin-splitting method.

\paragraph{Metrics.} Describe the metrics used for comparing methods.

\subsection{Synthetic Example 1: Uncertainty Quantification}

In general, there is no a single unchallenged aggreement about the
correct feature effect; each method models it in a slightly different
way. Therefore, for creating a universal evaluation framework we
reside on variations of the simple piecewise-linear function, where in
most cases, we can have an homophony about the correct effect.

In all the experiments of the example, we will use the following
generative distribution: \(p(\mathbf{x}) = p(x_3)p(x_2|x_1)p(x_1)\)
where \(x_1 \sim \mathcal{U}(0,1)\),
\(x_2 \sim \mathcal{N}(x_1, \sigma_2^2)\),
\(x_3 \sim \mathcal{N}(0, \sigma_3^2)\). 
%
The predictive function is:

\begin{equation}
  \label{eq:2}
  f(\mathbf{x}) = \begin{cases}
                    f_{\mathtt{lin}} + \alpha f_{\mathtt{int}} & \text{if $f_{\mathtt{lin}} < 0.5$ }\\
                    0.5 - f_{\mathtt{lin}} + \alpha f_{\mathtt{int}} & \text{if $0.5 \leq f_{\mathtt{lin}} < 1$}\\
                    \alpha f_{\mathtt{int}} &\text{otherwise}
                  \end{cases}
\end{equation}
%
where \(f_{\mathtt{lin}}(\mathbf{x}) = a_1 x_1 + a_2 x_2\) and
\(f_{inter}(\mathbf{x}) = x_1x_3\). The term \(f_{\mathtt{lin}}\) is a
linear function with \(a_1, a_2\) the weights of the corresponding
variables. The term \(f_{\mathtt{lin}}\) is the interaction between
the independent variables \(x_1\) and \(x_3\), with \(\alpha\) setting
the weight of the interaction term.

\subsubsection{Effect Without Uncertainty}

We test the feature effect when no interaction term is apparent,
i.e. \(\alpha=0\), there is no uncertainty about the value of \(x_2\),
i.e. \(x_2 = x_1\) and the weights are equal weights \(a_1=a_2=1\).
In this case, ground truth effect for \(f^{\mathtt{GT}}(x_1)\) is
\(x_1\) when \(0 \leq x_1 < 0.25\), \(-x_1\) when
\(0.25 \leq x_1 < 0.5\) and zero otherwise. In each position \(x_1\),
there is complete absence of heterogeneous effects, and therefore, the
uncertainty of the global explanation is zero,
i.e. \(f^{\mathcal{GT}}_{\sigma^2}(x_1) = 0\). In figure we present,
the effect of each method;

\begin{figure}[h]
  \centering
  \includegraphics[width=.23\textwidth]{example_3/dale_auto_bins.png}
  \includegraphics[width=.23\textwidth]{example_3/pdp_ice.png}
  \caption{Figure 1}
  \label{fig:ex-synth-1-1}
\end{figure}

\subsubsection{Effect With Uncertainty}

\subsection{Synthetic Example 2: Bin-Splitting}

The perfect bin-splitting according to the definition of the
minimization problem, is the one that creates:

\begin{itemize}
\item the biggest bins in regions with constant effect
\item the smallest bins with \(N_k > \tau\), in regions with non-constant effect
\end{itemize}

Prove that (a) this is a good objective, i.e., prove that on average,
this splitting gives the minimum distance from the ground truth ALE
and (b) that DP finds the optimal.

Set-ups:

For limited instances, measure the mean error and std (many seeds)
between gt ALE with uncertainty and approximation with bin-splitting
using (a) fixed-size bins (b) variable-size bins for many \(K\)

for many different \(N_k\)

\begin{itemize}
\item In case where the perfect bin-splitting is served by
  equally-sized bins, DP finds it automatically. Without DP, the user
  should try many different \(K\). and compare them the user's overhead of
  testing many differ
\end{itemize}

For many model-agnostic interpretation techniques
– including interaction detection methods – ground
truth information is usually not available on real-world
data. Therefore, well-constructed simulation experi-
ments with a known ground truth are often used for
empirical evaluations and comparisons, while only one
or few real-world datasets are used to demonstrate
practical applicability (e.g., see Friedman et al. (2008),
Fisher et al. (2019), Goldstein et al. (2015), Greenwell
et al. (2018), or Aas et al. (2021)). Hence, we follow
this commonly used approach to evaluate our method
using various simulation settings.
